Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2019-07-12 GWLDCI.tex
changes discussed on the phone

Meanwhile, there is a separate body of work studying SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv, Britto:2005fq, Cachazo:2004kj}. These integrals are calculated via Wilson loops. As noted above, the \hlfix{first author and Amat}{At other points, we say "some of us" instead of specifying like this. Choose a convention}\note{Response by Karen: I am the cause of the ``some of us'' and similar; it is the kind of phrasing I usually use, and I like it, but I am not strongly attached to it.} uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that are realized as an element of the positive Grassmannian $\Gr(k,n)$ \cite{wilsonloop}. In (separate) papers with the second author and Marcott, this work was extended to study these Feynman integrals in terms of positroid cells and the positive Grassmannians \cite{casestudy, non-orientable}. Others have tried different approaches to define a geometric space associated to these integrals in a manner similar to the Amplituhedron \cite{Amplituhedronsquared, HeslopStewart}.

@@

and thus speak of a propagator $p = (i, j)$ as being supported by the $i$th and $j$th edges. \sanote{moved/altered dfn and sandwiching paragrahs to improve flow}

@@

\begin{rmk}\label{alt F(P) rmk}
An equivalent definition of $F(P)$ is \bas F(P)  = \big(V(P) \setminus V(P^c) \big) \cup \{\text{all non-supporting vertices in $W$}\} \;,\eas 
i.e. $F(P)$ can be thought of as the set of vertices that either only support propagators in $P$ or do not support any propagators at all. Furthermore, note that by construction we have $\Prop(F(P)) \subseteq P$.  This containment is strict if there is any propagator $p$ in $P$ with the property that every vertex in the support of $p$ also supports a propagator not in $P$. \sanote{whoever added this last sentence, thank you. It is an oft overlooked important point.}
\end{rmk}

@@

In other words, a Wilson loop diagram, $(\cP, [n])$ is admissible only if $n \geq |\cP| +4$, and has no crossing propagators, no propagators of the form $p = (i, i+1)$, and no pairs of propagators that start and end on the same pair of non-adjacent edges.\note{Sian: can we leave this sentence out? By dropping ``iff'' to ``only if'' we make it sound like the converse isn't true.} \sanote{Lemma \ref{tausimpleplanarlem} uses the no propagators starting and stopping at same edge}

Note that if we take any admissible Wilson loop diagram and remove the non-supporting vertices, we obtain a weakly admissible Wilson loop diagram that may or may not be admissible itself.

@@

\begin{dfn}\label{equivdfn} 
Let $W = (\cP,[n])$ and $W' = (\cP',[n])$ be two admissible subdiagrams. We say that $W$ and $W'$ are related if
\begin{enumerate}
\item There exist exact subdiagrams $(P, V(P)) \subseteq W$ and $(P', V(P')) \subseteq W'$ with the same number of propagators and suppored on the same vertices; $|P| = |P'|$ and  $V(P) =  V(P')$.
\item The complementary subdiagrams (i.e. all other propagators and vertices) are identical, i.e. $(\cP \setminus P, [n]) = (\cP' \setminus P', [n])$. \sanote{made some changes} \note{I don't want to change it in case I've just misunderstood, but it seems to me that the vertex sets for the complementary subdiagrams should be $[n]$ not $V(P)^c$ and $V(P')^c$.} \sanote{good call. I had misread it to $V(P^c)$ and $V(P{'c}$, which is didn't say, and added some text. Fixed now.}
\end{enumerate}
Let $\sim$ denote the transitive closure of this relation. This is an equivalence relation on the set of all admissible Wilson loop diagrams with vertex set $[n]$.
\end{dfn}

@@

We have $W_1 \sim W_2$ because they are identical except on the exact subdiagrams 
\[(\{r, s\}, \{5,6,7,8,1\}) \text{ and }(\{u, v\}, \{5,6,7,8,1\})\] respectively. We also have $W_2 \sim W_3$ due to the exact subdiagrams 
\[(\{p, q\}, \{1,2,3,4,5\}) \text{ and }(\{a, b\}, \{1,2,3,4,5\}).\] 
However, the difference between $W_1$ and $W_3$ is the union of two exact subdiagrams, which is not itself exact. \sanote{labeled propagators to make easier to read} \end{eg}

As described in the introduction, Wilson loop diagrams come from the study of SYM $N = 4$ in the physics literature \st{and therefore} \hlfix{where each propagator corresponds to a section of  a $k$ dimesional vector bundle over twistor space. Thus}{I feel the strike out was too vague. The highlighted may be too technical. Can someone find baby bear's text?} each diagram $W$ is associated to a matrix of data $C(W)$, which we now describe.

@@

Throughout, we take the {\em matroid defined by $W$} to be the matroid $M(W)$ of Theorem \ref{thm WLD defines matroid}. Note from Definition~\ref{def:CWmatrix} that the vertices of the diagram $W$ correspond to columns of the associated matrix $C(W)$. Therefore, for any subset of vertices $V \subseteq [n]$, we can consider the restriction $C(W)|_V$ of $C(W)$ consisting of the columns corresponding to $V$. Since the non-zero entries of $C(W)$ are independent real variables, any subset $V$ that has fewer than $|V|$ rows of non-zero entries does not have full rank. This is equivalent to saying that if $V$ is a set of columns of $C(W)$ with $|\Prop(V)| < |V|$, $V$ is a dependent set. Any set subset of columns containing such a $V$ cannot be independent. Comparing this to Theorem \ref{thm:WLDmatroid}, we see that $M(W)$ is also the matroid realized by $C(W)$. \sanote{fixed dependent set argument}

@@

To prove (4), we need to show that $F(P)$ is maximally dependent. If $F(P) = [n]$ then this is automatic, so suppose not and let $v \in [n] \setminus F(P)$. In other words, $v \in V(P^c)$ and so $v$ supports some propagator $q \not\in P$.  Let $S \subseteq F(P)$ be an independent set of maximal size. Then $\Prop(S) \subseteq P$ (from the definition of $F(P)$) and no subset of $S$ supports fewer propagators than the number of vertices it contains (by Theorem~\ref{thm WLD defines matroid}) \st{this is one characterization of an independent set in $M(W)$)}. Since $v$ supports a new propagator $q \not\in P$, the set $S \cup  \subseteq F(P) \cup v$ also satisfies this independence condition. Furthermore, since $S \cup v$ is independent, by the first point in this lemma, $|S| = \rk(S) = \rk F(P)$. Similarly $|S \cup v| = |S|+1 = \rk (S \cup v) \leq \rk(F(P) \cup v)$.  Thus $\rk(F(P) \cup\{v\}) = \rk(F(P)) + 1$, as required. \sanote{tweaked}
\end{proof}

@@

In this section, we take this analysis further, showing the converse of Theroem [ref] from loc. cit. (see Theorem \ref{same matroid iff equiv}. We also enumerate the number of diagrams that map to an particular matroid, Corollary \ref{number of equiv diagrams}. To do this, we establish a connectection between Wilson loop diagrams and chord diagrams, subsection \ref{sec: polygon partitions}, and prove some matroidal properties of exact subdiagrams in subsection \ref{sec: exact diagram matroidal props}  including that they correspond to uniform matroids (Theorem \ref{exactuniformthm}), and the associated matroid can be written as as a contraction by the complementary propagator flat (Theorem \ref{exact diagrams contractions}).\sanote{need to discuss positroids, and whether positroids should appear earlier so that they can be discussed here.}

@@

\begin{dfn}\label{WLDtriangulationdfn}
  Let $W = (\cP, [n])$ be an admissible Wilson loop diagram.  The \emph{polygon dissection} associated to $W$, denoted $\tau(W)$, is defined as follows.
  \begin{itemize}
  \item The vertices of $\tau(W)$ correspond to the edges of $W$.
  \item Labeling the vertices of $\tau(W)$ with the edge number of $W$gives a cyclic order to the vertices. Connecting consecutive vertices gives a graph theoretic cycle called the polygon of $\tau(W)$.
  \item Each propagator of $W$ defines a chord \note{We were originally thinking in terms of chord diagrams hence the language ``chord'' but I think ``diagonal'' would be more natural now.  Any thoughts against before I change it?}\sanote{go for it.} edge of $\tau(W)$; specifically,  a propagator $(i,j) \in \cP$ defines a chord connecting the vertices $i$ and $j$ in $\tau(W)$.
  \end{itemize}
\end{dfn}

@@

\begin{dfn}
  Let $W$ be an admissible Wilson loop diagram and $\tau(W)$ its polygon dissection. A \emph{triangulated piece} of $\tau(W)$ is a 2-connected subgraph of $\tau(W)$ which is a triangulation. We will take the convention that a subgraph consisting of a single chord edge is called a \emph{trivial} triangulated piece.
A {\em maximal} triangulated piece is one which is not contained in any strictly larger triangulated piece.
\end{dfn}

@@

Next, we introduce the idea of retriangulations. 
\begin{dfn} \label{retriangulation}
  If $W = (\cP, [n])$ and $W' = (\cP', [n])$ are weakly admissible Wilson loop diagrams, then $\tau(W)$ and $\tau(W')$  differ by a \emph{retriangulation} if there is a bijection $f$ between the maximal triangulated pieces of $\tau(W)$ and the maximal triangulated pieces of $\tau(W')$ such that for any maximal triangulated piece $t$ of $\tau(W)$ the vertex sets of $t$ and $f(t)$ are the same.
\end{dfn}
More casually this definition says that the maximal triangulated pieces of $\tau(W)$ and $\tau(W')$ are the same (so far as their vertex sets are concerned) but that some or all of the triangulated pieces may be differently triangulated while the region being triangulated by each piece (which is defined by the vertex set) remains unchanged. This defintion of retriangulation makes sense for any two polygon dissections of the same polygon, not just for those in the image of $\tau$. \sanote{rearranged the paras around dfn \ref{retriangulation} for flow}

@@

\begin{eg}\label{eg:subtrees}
Consider the following Wilson loop diagram $W$ and its corresponding chord diagram $\tau(W)$. 
\bas W = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{10}{1.5}
	\drawnumbers
	\drawprop{1}{0}{4}{0}
	\drawprop{2}{0}{4}{-1}
	\drawprop{4}{1}{8}{0}
	\drawprop{8}{1}{1}{-1}
	\drawprop{5}{0}{7}{0}
%\draw (0,-2) --(0,2); horiontal
%\draw (-2,0) --(2,0);
%\draw (0,-2) node {\footnotesize $-y$};
%\draw (-2,0) node {\footnotesize $-x$};
\draw (-.3,1.) node {\footnotesize $p$};
\draw (-.2,.65) node {\footnotesize $q$};
\draw (0,-.4) node {\footnotesize $r$};
\draw (.7, 0) node {\footnotesize $s$};
		\end{scope}
	\end{tikzpicture}  \; \quad
\tau(W) = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawpolypart{10}{1.5}
    \drawnumbersshift
    \drawchord{1}{4}
    \drawchord{2}{4}
    \drawchord{4}{8}
    \drawchord{8}{1}
    \drawchord{5}{7}
\draw (-.3,1.) node {\footnotesize $p$};
\draw (-.2,.65) node {\footnotesize $q$};
\draw (0,-.4) node {\footnotesize $r$};
\draw (.7, 0) node {\footnotesize $s$};
	\end{scope}
	\end{tikzpicture} \eas

@@

Each of these $m-3+j$ \hlfix{internal}{diagonal? chord?} edges corresponds to a propagator in $W$; call this set of propagators $P$.  Next we count the size of $V(P)$.  Each of the $m$ vertices in the outer face of $t$ corresponds to an edge of $W$. These $m$ edges define $j$ connected components of the outer polygon of $W$. Thus the set $V(P)$ has $m+j$ vertices.  In other words, 
\[|V(P)| = m+j = |P| +3\;.\]
Thus the subdiagram $(P,V(P))$ defined by $t$ is exact.

@@

Now suppose $|P|>1$.  Let $j=|P|$, $m= j+3=|V(P)|$, and 
suppose that the set $V(P)$ defines $c$ disjoint cyclic intervals of $[n]$. Then $t$ has $m-c$ vertices.  If $t$ were a triangulation, $t$ would have $j -c$ \hlfix{internal}{here, internal means that it doesn't define the outer face of $t$ in the induced embedding. Different use of the word than above?} edges.

The graph $t$ has $j$ edges that come from propagators, and $m - 2c$ edges that come from the boundary polygon of $\tau(W)$. Since $t$ has $m-c$ vertices, it has $m-c$ \hlfix{external}{means defines outer face, not just and outer edge of $\tau(W)$} edges, of which $c$ come from propagators. Therefore, of the $j$ edges of $t$ that come from propagators, $j-c$ are internal to the connected component. Therefore, $t$ is a triangulated piece.

@@

\begin{proof}
Combining Lemmas \ref{decompositionlem} and \ref{lem triang to exact} yields the unique decomposition into maximal exact subdiagrams, and Corollary \ref{maxtriangdisjointcor} ensures that no propagator appears in more than one subdiagram in this decomposition. Since the chord edges of $\tau(W)$ correspond to the propagators of $W$, the decomposition of $\tau(W)$ induces a partition of $\cP$.
\end{proof}

@@

We begin by proving some useful facts about propagator flats, and flats of matroids associated to admissible Wilson loop diagrams more generally. Recall from Definition~\ref{VPropdfn} that the propagator flat $F(P)$ of $P \subseteq \cP$ denotes the set $V(P^c)^c$ of vertices that \st{do not support any propagators not in $P$.}\hlfix{only support propagators in $P$ (or no propagators)}{too many negatives. Can we use this instead?}

@@

\begin{lem} \label{maxexactcomplementrank}
Let $W = (\cP, [n])$ be an admissible Wilson loop diagram, and $P \subseteq \cP$. Then: \begin{enumerate}
\item If $(P,V(P))$ is an exact subdiagram of $W$, then $F(P^c)$ is a dependent flat in $W$. \note{Should we emphasize that all these $P^c$ are wrt $W$ not the subdiagram?} \sanote{yes. done}
\item If $(P,V(P))$ is a maximal exact subdiagram of $W$, then $F(P^c)$ is a cyclic flat in $W$.
\item If $(P,V(P))$ is an exact subdiagram in $W$, then $\rk(F(P^c)) = |P^c|$.
\end{enumerate} \sanote{end here}
\end{lem}

@@

In this section we have characterized the correspondence between Wilson loop diagrams and positroids, showing that Wilson Loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}). Corollary \ref{number of equiv diagrams} enumerates the fibres\note{Is ``enumerates the fibres'' correct?} of the map from Wilson Loop diagrams to positroids.  The obvious next question would be to enumerate the number of equivlence classes.

@@

The map $\tau$ defined in Definition \ref{WLDtriangulationdfn} assigns a dissection of an $n$-gon to a Wilson loop diagram $W = (\cP, [n])$. Recall that if $|\cP| = k$, then $\tau(W)$ is a dissection with $k$ chords. Therefore, the map $\tau$ associates to each weakly admissible Wilson loop diagram an $n-3 - k$ dimensional face of the associahedron $A_n$. This is illustrated for $n=5$ on the left in Figure~\ref{associahedron5gon}, and with some dissections illustrated for $n=6$ on the right.

>>>>

Meanwhile, there is a separate body of work studying SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv, Britto:2005fq, Cachazo:2004kj}. These integrals are calculated via Wilson loops. As noted above, Agarwala and Marin-Amat uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that are realized as an element of the positive Grassmannian $\Gr(k,n)$ \cite{wilsonloop}. In (separate) papers with the second author and Marcott, this work was extended to study these Feynman integrals in terms of positroid cells and the positive Grassmannians \cite{casestudy, non-orientable}. Others have tried different approaches to define a geometric space associated to these integrals in a manner similar to the Amplituhedron \cite{Amplituhedronsquared, HeslopStewart}.

@@

Thus we may speak of a propagator $p = (i, j)$ as being supported by the $i$th and $j$th edges.

@@

\begin{rmk}\label{alt F(P) rmk}
An equivalent definition of $F(P)$ is \bas F(P)  = \big(V(P) \setminus V(P^c) \big) \cup \{\text{all non-supporting vertices in $W$}\} \;,\eas 
i.e. $F(P)$ can be thought of as the set of vertices that either only support propagators in $P$ or do not support any propagators at all. Furthermore, note that by construction we have $\Prop(F(P)) \subseteq P$.  This containment is strict if there is any propagator $p$ in $P$ with the property that every vertex in the support of $p$ also supports a propagator not in $P$. 
\end{rmk}

@@

There are a few things to note here.  First, what we call weakly admissible is called admissible in \cite{wilsonloop}.  Many of the results below apply when the hypothesis of admissibility is changed to weak admissibility.

Note also that if two propagators have the same supporting edges, or if a propagator is supported on adjacent edges then the Wilson loop diagram is not admissible.

Further if we take any admissible Wilson loop diagram and remove the non-supporting vertices, we obtain a weakly admissible Wilson loop diagram that may or may not be admissible itself.

@@

\begin{dfn}\label{equivdfn} 
Let $W = (\cP,[n])$ and $W' = (\cP',[n])$ be two admissible subdiagrams. We say that $W$ and $W'$ are related if
\begin{enumerate}
\item There exist exact subdiagrams $(P, V(P)) \subseteq W$ and $(P', V(P')) \subseteq W'$ with the same number of propagators and suppored on the same vertices; $|P| = |P'|$ and  $V(P) =  V(P')$.
\item The complementary subdiagrams (i.e. all other propagators and vertices) are identical, i.e. $(\cP \setminus P, [n]) = (\cP' \setminus P', [n])$.
\end{enumerate}
Let $\sim$ denote the transitive closure of this relation. This is an equivalence relation on the set of all admissible Wilson loop diagrams with vertex set $[n]$.
\end{dfn}

@@

We have $W_1 \sim W_2$ because they are identical except on the exact subdiagrams 
\[(\{r, s\}, \{5,6,7,8,1\}) \text{ and }(\{u, v\}, \{5,6,7,8,1\})\] respectively. We also have $W_2 \sim W_3$ due to the exact subdiagrams 
\[(\{p, q\}, \{1,2,3,4,5\}) \text{ and }(\{a, b\}, \{1,2,3,4,5\}).\] 
However, the difference between $W_1$ and $W_3$ is the union of two exact subdiagrams, which is not itself exact.

As described in the introduction, Wilson loop diagrams come from the study of SYM $N = 4$ in the physics literature and in particular, see \cite{Adamo:2012xe} for details, this gives that each diagram $W$ is associated to a matrix of data $C(W)$, which we now describe.

@@

Throughout, we take the {\em matroid defined by $W$} to be the matroid $M(W)$ of Theorem \ref{thm WLD defines matroid}. Note from Definition~\ref{def:CWmatrix} that the vertices of the diagram $W$ correspond to columns of the associated matrix $C(W)$. Therefore, for any subset of vertices $V \subseteq [n]$, we can consider the restriction $C(W)|_V$ of $C(W)$ consisting of the columns corresponding to $V$. Since the non-zero entries of $C(W)$ are independent real variables, any subset $V$ that has fewer than $|V|$ rows of non-zero entries does not have full rank. This is equivalent to saying that if $V$ is a set of columns of $C(W)$ with $|\Prop(V)| < |V|$, $V$ is a dependent set. Any set subset of columns containing such a $V$ cannot be independent. Comparing this to Theorem \ref{thm:WLDmatroid}, we see that $M(W)$ is also the matroid realized by $C(W)$.

@@

To prove (4), we need to show that $F(P)$ is maximally dependent. If $F(P) = [n]$ then this is automatic, so suppose not and let $v \in [n] \setminus F(P)$. In other words, $v \in V(P^c)$ and so $v$ supports some propagator $q \not\in P$.  Let $S \subseteq F(P)$ be an independent set of maximal size. Then $\Prop(S) \subseteq P$ (from the definition of $F(P)$) and no subset of $S$ supports fewer propagators than the number of vertices it contains (by Theorem~\ref{thm WLD defines matroid}). Since $v$ supports a new propagator $q \not\in P$, the set $S \cup \{v\} \subseteq F(P) \cup \{v\}$ also satisfies this independence condition. Furthermore, since $S \cup \{v\}$ is independent, by the first point in this lemma, $|S| = \rk(S) = \rk F(P)$. Similarly $|S \cup \{v\}| = |S|+1 = \rk (S \cup \{v\}) \leq \rk(F(P) \cup \{v\})$.  Thus $\rk(F(P) \cup\{v\}) = \rk(F(P)) + 1$, as required.
\end{proof}

@@

In this section, we take this analysis further, showing the converse of Theroem [ref] from loc. cit. (see Theorem \ref{same matroid iff equiv}. We also enumerate the number of diagrams that map to an particular matroid, Corollary \ref{number of equiv diagrams}. To do this, we establish a connectection between Wilson loop diagrams and polygon dissections, subsection \ref{sec: polygon partitions}, and prove some matroidal properties of exact subdiagrams in subsection \ref{sec: exact diagram matroidal props}  including that they correspond to uniform matroids (Theorem \ref{exactuniformthm}), and the associated matroid can be written as as a contraction by the complementary propagator flat (Theorem \ref{exact diagrams contractions}).\sanote{need to discuss positroids, and whether positroids should appear earlier so that they can be discussed here.}

@@

\begin{dfn}\label{WLDtriangulationdfn}
  Let $W = (\cP, [n])$ be an admissible Wilson loop diagram.  The \emph{polygon dissection} associated to $W$, denoted $\tau(W)$, is defined as follows.
  \begin{itemize}
  \item The vertices of $\tau(W)$ correspond to the edges of $W$.
  \item Labeling the vertices of $\tau(W)$ with the edge number of $W$gives a cyclic order to the vertices. Connecting consecutive vertices gives a graph theoretic cycle called the polygon of $\tau(W)$.
  \item Each propagator of $W$ defines a diagonal edge of $\tau(W)$; specifically,  a propagator $(i,j) \in \cP$ defines a diagonal connecting the vertices $i$ and $j$ in $\tau(W)$.
  \end{itemize}
\end{dfn}

@@

\begin{dfn}
  Let $W$ be an admissible Wilson loop diagram and $\tau(W)$ its polygon dissection. A \emph{triangulated piece} of $\tau(W)$ is a 2-connected subgraph of $\tau(W)$ which is a triangulation. We will take the convention that a subgraph consisting of a single diagonal edge is called a \emph{trivial} triangulated piece.
A {\em maximal} triangulated piece is one which is not contained in any strictly larger triangulated piece.
\end{dfn}

@@

Next, we introduce the idea of retriangulations. 
\begin{dfn} \label{retriangulation}
  If $W = (\cP, [n])$ and $W' = (\cP', [n])$ are weakly admissible Wilson loop diagrams, then $\tau(W)$ and $\tau(W')$  differ by a \emph{retriangulation} if there is a bijection $f$ between the maximal triangulated pieces of $\tau(W)$ and the maximal triangulated pieces of $\tau(W')$ such that for any maximal triangulated piece $t$ of $\tau(W)$ the vertex sets of $t$ and $f(t)$ are the same.
\end{dfn}
More casually this definition says that the maximal triangulated pieces of $\tau(W)$ and $\tau(W')$ are the same (so far as their vertex sets are concerned) but that some or all of the triangulated pieces may be differently triangulated while the region being triangulated by each piece (which is defined by the vertex set) remains unchanged. This defintion of retriangulation makes sense for any two polygon dissections of the same polygon, not just for those in the image of $\tau$.

@@

\begin{eg}\label{eg:subtrees}
Consider the following Wilson loop diagram $W$ and its corresponding polygon dissection $\tau(W)$. 
\bas W = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{10}{1.5}
	\drawnumbers
	\drawprop{1}{0}{4}{0}
	\drawprop{2}{0}{4}{-1}
	\drawprop{4}{1}{8}{0}
	\drawprop{8}{1}{1}{-1}
	\drawprop{5}{0}{7}{0}
%\draw (0,-2) --(0,2); horiontal
%\draw (-2,0) --(2,0);
%\draw (0,-2) node {\footnotesize $-y$};
%\draw (-2,0) node {\footnotesize $-x$};
\draw (-.3,1.) node {\footnotesize $p$};
\draw (-.2,.65) node {\footnotesize $q$};
\draw (0,-.4) node {\footnotesize $r$};
\draw (.7, 0) node {\footnotesize $s$};
		\end{scope}
	\end{tikzpicture}  \; \quad
\tau(W) = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawpolypart{10}{1.5}
    \drawnumbersshift
    \drawchord{1}{4}
    \drawchord{2}{4}
    \drawchord{4}{8}
    \drawchord{8}{1}
    \drawchord{5}{7}
\draw (-.3,1.) node {\footnotesize $p$};
\draw (-.2,.65) node {\footnotesize $q$};
\draw (0,-.4) node {\footnotesize $r$};
\draw (.7, 0) node {\footnotesize $s$};
	\end{scope}
	\end{tikzpicture} \eas

@@

Each of these $m-3+j$ edges corresponds to a propagator in $W$; call this set of propagators $P$.  Next we count the size of $V(P)$.  Each of the $m$ vertices in the outer face of $t$ corresponds to an edge of $W$. These $m$ edges define $j$ connected components of the outer polygon of $W$. Thus the set $V(P)$ has $m+j$ vertices.  In other words, 
\[|V(P)| = m+j = |P| +3\;.\]
Thus the subdiagram $(P,V(P))$ defined by $t$ is exact.

@@

Now suppose $|P|>1$.  Let $j=|P|$, $m= j+3=|V(P)|$, and 
suppose that the set $V(P)$ defines $c$ disjoint cyclic intervals of $[n]$. Then $t$ has $m-c$ vertices.  If $t$ were a triangulation, $t$ would have $j -c$ internal edges.

The graph $t$ has $j$ edges that come from propagators, and $m - 2c$ edges that come from the boundary polygon of $\tau(W)$. Since $t$ has $m-c$ vertices, it has $m-c$ external edges, of which $c$ come from propagators. Therefore, of the $j$ edges of $t$ that come from propagators, $j-c$ are internal to the connected component. Therefore, $t$ is a triangulated piece.

@@

\begin{proof}
Combining Lemmas \ref{decompositionlem} and \ref{lem triang to exact} yields the unique decomposition into maximal exact subdiagrams, and Corollary \ref{maxtriangdisjointcor} ensures that no propagator appears in more than one subdiagram in this decomposition. Since the diagonal edges of $\tau(W)$ correspond to the propagators of $W$, the decomposition of $\tau(W)$ induces a partition of $\cP$.
\end{proof}

@@

We begin by proving some useful facts about propagator flats, and flats of matroids associated to admissible Wilson loop diagrams more generally. Recall from Definition~\ref{VPropdfn} that the propagator flat $F(P)$ of $P \subseteq \cP$ denotes the set $V(P^c)^c$ of vertices that only support propagators in $P$ (or no propagators).

@@

\begin{lem} \label{maxexactcomplementrank}
Let $W = (\cP, [n])$ be an admissible Wilson loop diagram, and $P \subseteq \cP$. Then: \begin{enumerate}
\item If $(P,V(P))$ is an exact subdiagram of $W$, then $F(P^c)$ is a dependent flat in $W$.
\item If $(P,V(P))$ is a maximal exact subdiagram of $W$, then $F(P^c)$ is a cyclic flat in $W$.
\item If $(P,V(P))$ is an exact subdiagram in $W$, then $\rk(F(P^c)) = |P^c|$.
\end{enumerate} \sanote{end here}
\end{lem}

@@

In this section we have characterized the correspondence between Wilson loop diagrams and positroids, showing that Wilson Loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}). Corollary \ref{number of equiv diagrams} enumerates the elements of each fibre of the map from Wilson Loop diagrams to positroids.  The obvious next question would be to enumerate the number of equivlence classes.

@@

The map $\tau$ defined in Definition \ref{WLDtriangulationdfn} assigns a dissection of an $n$-gon to a Wilson loop diagram $W = (\cP, [n])$. Recall that if $|\cP| = k$, then $\tau(W)$ is a dissection with $k$ diagonals. Therefore, the map $\tau$ associates to each weakly admissible Wilson loop diagram an $n-3 - k$ dimensional face of the associahedron $A_n$. This is illustrated for $n=5$ on the left in Figure~\ref{associahedron5gon}, and with some dissections illustrated for $n=6$ on the right.
